---
title: "Quantifying Spatiotemoral Patterns in Small Cell Carbon Uptake"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# A file focused on quantifying, visualizing and identifying statistical trends in small-cell marine carbon uptake. Using data from 13-C and # 15-N labeled isotopic field experiments. 


# Upload required libraries:
```{r}

suppressMessages(library(ggplot2))
suppressMessages(library(caret))
suppressMessages(library(ggpubr))
suppressMessages(library(dplyr))
suppressMessages(library(viridis))
suppressMessages(library(forcats))
suppressMessages(library(plotrix))
suppressMessages(library(ggsci))
suppressMessages(library(RColorBrewer))
suppressMessages(library(ggridges))
suppressMessages(library(PCAtools))
suppressMessages(library(ggfortify))
suppressMessages(library(plyr))
suppressMessages(library(plotrix))


```


# Load the data:
```{r}

data <- read.csv("Carbon_rates.csv")

#Turn your 'treatment' column into a character vector
data$Station <- as.character(data$Station)
#Then turn it back into a factor with the levels in the correct order
data$Station <- factor(data$Station, levels=unique(data$Station))

boxplot_data <- read.csv("Boxplot_rates.csv")


```


# Smoothed means: Visualize small cell contributions over spatial gradients.
```{r}


facet_plot <-  boxplot_data %>%

  
  mutate(Type = fct_relevel(Type, 
            "Carbon", "New production", "Nitrate", "Silica")) %>%
  
  ggplot(aes(x = DFS, y = Percent_contribution, color = Type, fill = Type)) +
    stat_smooth(aes(lty = Type), method = "loess", show.legend = TRUE, 
                alpha = 0.7, se = TRUE, level = 0.9, span = 0.9) +
    geom_rug(sides = 'b', alpha = 1, show.legend = FALSE, size = 0.3, color = 'black') +
    ylab("% Contribution of small cells (< 5µm)") +
    xlab("Distance from shore (km)") +
    theme_bw() +
    ylim(0, 100) +
    
    # Optional color schemes:
    #scale_color_lancet() +
    #scale_fill_lancet() +
    #scale_color_brewer(palette = "Dark")
    #scale_fill_brewer(palette = "Dark")
  
    scale_color_manual(values=c("red", "black", "#E69F00", "#56B4E9")) +
    scale_fill_manual(values=c("red", "black", "#E69F00", "#56B4E9")) +
    theme(legend.key=element_blank(), legend.position = "top", legend.title = element_blank())

facet_plot


Boxplot_SCC <- boxplot_data %>%
  
    mutate(Regions = fct_relevel(Region3, 
            "Coastal", "Transition", "HNLC")) %>%
  
  
    ggplot(aes(x = Regions, y = Percent_contribution, fill = log(Total_uptake))) +
    geom_boxplot(alpha=0.5, outlier.shape = NA, position = dodge) +
    geom_point(pch = 21, position = position_jitterdodge()) +
    theme(legend.position="top") +
    facet_wrap(~Type) +
    xlab("") +
    ylab("") +
    scale_fill_viridis(option = 'H') +
    #scale_fill_discrete(labels = c("Small (<5µm)", "Large (>5µm)")) +
    theme_bw() +
    theme(legend.key=element_blank(), legend.position = "top", legend.title = element_blank())

Boxplot_SCC



pdf(file = "smoothaverage_plot.pdf",  
    width = 12, 
    height = 8) 

facet_plot

dev.off()

```


# Total contributions: Visualize the magnitude of contributions between large and small cells over spatial gradients.
```{r}

# Stack the data:
stacked_data <- select(boxplot_data, c('Type', 'Region3', 'Small_uptake', 'Large_uptake', 'Season'))
stacked_dataframe <- na.omit(cbind(stacked_data[c(1, 2, 5)], stack(stacked_data[c(3, 4)])))


# Carbon
# Isolate Carbon:

C_stacked_df <- stacked_dataframe[stacked_dataframe$Type == "Carbon", ]
C_stacked_df$Region <- factor(C_stacked_df$Region3)
C_stacked_df$ind3 <- as.numeric(C_stacked_df$ind)

dodge <- position_dodge(width = 0.9)

#C_stacked_df$Regions <- as.numeric(C_stacked_df$Region) + 
  #3 * (as.numeric((C_stacked_df$ind)) - 0.2)

C_magnitude_boxplot <- C_stacked_df %>%
  

  mutate(Region = fct_relevel(Region, 
            "Coastal", "Transition", "HNLC")) %>%
  
  
  ggplot(aes(x = Region, y = values, fill = ind)) +
  geom_boxplot(alpha=0.5, outlier.shape = NA, position = dodge) +
  geom_point(pch = 21, position = position_jitterdodge()) +
  theme(legend.position="top") +
  xlab("") +
  ylab("") +
  ggtitle(expression("Carbon uptake (µg C L"^-1*"d"^-1*")")) +
  scale_y_sqrt(limits = c(0, 900), breaks = (0:30)^2) +
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9")) +
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9")) +
  scale_fill_discrete(labels = c("Small (<5µm)", "Large (>5µm)")) +
  theme_bw() +
  theme(legend.key=element_blank(), legend.position = "top", legend.title = element_blank())

C_magnitude_boxplot




# Nitrate
# Isolate Nitrate:

N_stacked_df <- stacked_dataframe[stacked_dataframe$Type == "Nitrate", ]
N_stacked_df$Region <- factor(N_stacked_df$Region3)
N_stacked_df$ind3 <- as.numeric(N_stacked_df$ind)


N_magnitude_boxplot <- N_stacked_df %>%
  

  mutate(Region = fct_relevel(Region, 
            "Coastal", "Transition", "HNLC")) %>%
  
  
  ggplot(aes(x = Region, y = values, fill = ind)) +
  geom_boxplot(alpha=0.5, outlier.shape = NA, position = dodge) +
  geom_point(pch = 21, position = position_jitterdodge()) +
  theme(legend.position="top") +
  xlab("") +
  ylab("") +
  ggtitle(expression("Nitrate uptake (µg N L"^-1*"d"^-1*")")) +
  scale_y_sqrt(limits = c(0, 144), breaks = (0:12)^2) +
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9")) +
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9")) +
  scale_fill_discrete(labels = c("Small (<5µm)", "Large (>5µm)")) +
  theme_bw() +
  theme(legend.key=element_blank(), legend.position = "top", legend.title = element_blank())

N_magnitude_boxplot


# New production
# Isolate NP:

NP_stacked_df <- stacked_dataframe[stacked_dataframe$Type == "New production", ]
NP_stacked_df$Region <- factor(NP_stacked_df$Region3)
NP_stacked_df$ind3 <- as.numeric(NP_stacked_df$ind)

dodge <- position_dodge(width = 0.9)

#NP_stacked_df$Region3 <- as.numeric(NP_stacked_df$Region3) + 
  #3 * (as.numeric((NP_stacked_df$ind)) - 0.2)

NP_magnitude_boxplot <- NP_stacked_df %>%
  

  mutate(Region = fct_relevel(Region, 
            "Coastal", "Transition", "HNLC")) %>%
  
  
  ggplot(aes(x = Region3, y = values, fill = ind)) +
  geom_boxplot(alpha=0.5, outlier.shape = NA, position = dodge) +
  geom_point(pch = 21, position = position_jitterdodge()) +
  theme(legend.position="top") +
  xlab("") +
  ylab("") +
  ggtitle(expression("New production (µg C L"^-1*"d"^-1*")")) +
  scale_y_sqrt(limits = c(0, 260), breaks = (0:16)^2) +
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9")) +
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9")) +
  scale_fill_discrete(labels = c("Small (<5µm)", "Large (>5µm)")) +
  theme_bw() +
  theme(legend.key=element_blank(), legend.position = "top", legend.title = element_blank())

NP_magnitude_boxplot


# Silica
# Isolate Si:

Si_stacked_df <- stacked_dataframe[stacked_dataframe$Type == "Silica", ]
Si_stacked_df$Region <- factor(Si_stacked_df$Region3)
Si_stacked_df$ind3 <- as.numeric(Si_stacked_df$ind)

Si_magnitude_boxplot <- Si_stacked_df %>%
  

  mutate(Region = fct_relevel(Region, 
            "Coastal", "Transition", "HNLC")) %>%
  
  
  ggplot(aes(x = Region3, y = values, fill = ind)) +
  geom_boxplot(alpha=0.5, outlier.shape = NA, position = dodge) +
  geom_point(pch = 21, position = position_jitterdodge()) +
  theme(legend.position="top") +
  xlab("") +
  ylab("") +
  ggtitle(expression("Silica Uptake (µmol Si L"^-1*"d"^-1*")")) +
  scale_y_sqrt(limits = c(0, 16), breaks = (0:4)^2) +
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9")) +
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9")) +
  scale_fill_discrete(labels = c("Small (<5µm)", "Large (>5µm)")) +
  theme_bw() +
  theme(legend.key=element_blank(), legend.position = "top", legend.title = element_blank())

Si_magnitude_boxplot


fig <- ggarrange(C_magnitude_boxplot, N_magnitude_boxplot, 
                 NP_magnitude_boxplot, Si_magnitude_boxplot, common.legend = TRUE)


pdf(file = "Mag_boxplots.pdf", 
    width = 12, 
    height = 10)

fig

dev.off()

```


# Particulate elemental ratios:
```{r}

# Select an dstack relevant data:
ridges_data <- select(data, c('Region3','Small_POC_PON', 'Large_POC_PON', 'POC_PON'))
ridges_data <- ridges_data[!is.na(as.numeric(ridges_data$Large_POC_PON)), ]

stacked_ridges <- cbind(ridges_data[c(1)], stack(ridges_data[c(2, 3)]))
stacked_ridges$RR <- as.numeric(stacked_ridges$values)



stacked_ridges <- na.omit(stacked_ridges)


ridges_plot <- stacked_ridges %>%
  

  mutate(Region = fct_relevel(Region3, 
            "Coastal", "Transition", "HNLC")) %>%
  
  ggplot(aes(x = RR, y = Region3, group = interaction(ind, Region3), fill = ind)) +
    stat_density_ridges(aes(fill = ind), quantile_lines = TRUE, quantiles = 2, alpha = 0.6, scale = 1, 
                        jittered_points = FALSE, point_shape = '|', rel_min_height = 0.01) +
    theme_bw() +
    ylab("") +
    scale_x_continuous(name="Redfield Ratio", breaks=c(0, 5, 6.6, 10, 20, 30)) +
    scale_color_manual(values=c("#E69F00", "#56B4E9", "#999999")) +
    scale_fill_manual(values=c("#E69F00", "#56B4E9", "#999999")) +
    scale_fill_discrete(labels = c("Small (<5µm)", "Large (>5µm)")) +
    theme(legend.key=element_blank(), legend.position = "top", legend.title = element_blank())


ridges_plot


pdf(file = "Ridges_plot.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 6) # The height of the plot in inches

ridges_plot

dev.off()


```


# Create a function to compile averages, medians, standard errors and deviations:
```{r}


means <- function(data_variable){
  
  avg <- mean(as.numeric(data_variable))
  med <- median(as.numeric(data_variable))
  SD <- sd(as.numeric(data_variable))
  SE <- std.error(as.numeric(data_variable))

  output <- list(avg, med, SD, SE, colnames = colnames)
  names(output) <- c("average", "median", "standard deviation", "standard error")
  
  return(output)
  
}


# Calculate statistics:
means(ridges_data$Small_POC_PON)
means(ridges_data$Large_POC_PON)


```


Create a function to calculate regression statistics for all dependent and independent variables:
```{r}

# Get a data frame
data_df <- read.csv("Carbon_rates.csv")

# Specify a workable data frame:
dt = data.frame(data_df)



# Create a function that takes dependent variables as input, and returns regression summaries for each independent variable:

stats_summary <- function(y_string, dataframe){   # Need to input y_string as a string
  

  # Specify columns to regress:
  y_col = y_string
  x_col = c("Temp", "Salinity", "PAR", "MLD", "DFS", "Nitrate", "Silicate", "DO")  # Independent variables
  
  ss <- expand.grid(y=y_col, x=x_col, stringsAsFactors = F) %>%
    mutate(formula = paste(y,"~",x)) %>%
    group_by(formula) %>%
    mutate(r_sq = summary(lm(formula, data=dt))$r.squared) %>%
    mutate(p = summary(lm(formula, data=dt))$coefficients[2,4]) %>%
    mutate(coeff = summary(lm(formula, data=dt))$coefficients[2,1]) %>%
    ungroup()
  
  return(ss)
  
}

```


Apply the function to each measurement:
```{r}

# Optional: remove winter:
dt <- dt[!dt$Season == "Winter", ]


# Carbon:
stats_summary("Percent_Nano_Carbon", dt)
stats_summary("Carbon_Rate", dt)
stats_summary("Nano_Carbon_Rate", dt)


# Nitrogen:
stats_summary("Percent_Nano_Nitrate", dt)
stats_summary("Nitrate_Rate", dt)
stats_summary("Nano_Nitrate_Rate", dt)


# Silica:
stats_summary("Percent_Nano_Silica", dt)
stats_summary("Silica_Rate", dt)
stats_summary("Nano_Silica_Rate", dt)


# New production:
stats_summary("Percent_Nano_NP", dt)
stats_summary("NP_Rate", dt)
stats_summary("Nano_NP_Rate", dt)


```


Principal Component Analysis:
```{r}

# Re-read the dataframe:
data <- read.csv("Carbon_rates.csv")

# Create suitable dataframes:

# Totals and PSCC:
df_totals <- data %>% select(Silica_Rate, Carbon_Rate, Nitrate_Rate, NP_Rate, # Totals
                             Percent_Nano_Silica, Percent_Nano_Carbon, Percent_Nano_Nitrate, Percent_Nano_NP,  # Smalls
                             SST, MLD, Nitrate, DO, Chlorophyll) # Environmental

df_totals_cats <- data %>% select(Silica_Rate, Carbon_Rate, Nitrate_Rate, NP_Rate, # Totals
                             Percent_Nano_Silica, Percent_Nano_Carbon, Percent_Nano_Nitrate, Percent_Nano_NP,  # Smalls
                             SST, MLD, Nitrate, DO, Chlorophyll,  # Environmental
                             Season, Region, Region2, Region3)  # Cats

# Small and large rates:
df_sfs <- data %>% select(Micro_Silica_Rate, Micro_Carbon_Rate, Micro_Nitrate_Rate, Micro_NP_Rate, # Totals
                             Nano_Silica_Rate, Nano_Carbon_Rate, Nano_Nitrate_Rate, Nano_NP_Rate,  # Smalls
                             Temp, MLD, Nitrate, Silicate, DO, Chlorophyll) # Environmental

df_sfs_cats <- data %>% select(Micro_Silica_Rate, Micro_Carbon_Rate, Micro_Nitrate_Rate, Micro_NP_Rate, # Totals
                             Nano_Silica_Rate, Nano_Carbon_Rate, Nano_Nitrate_Rate, Nano_NP_Rate,  # Smalls
                             Temp, MLD, Nitrate, Silicate, DO, Chlorophyll,  # Environmental
                             Season, Region, Region2, Region3)  # Cats 



df_SCC <- data %>% select(Percent_Nano_Silica, Percent_Nano_Carbon, Percent_Nano_Nitrate, Percent_Nano_NP, 
                              Nitrate, Silicate, Temp, MLD, Sigma, PAR, DO, DFS, Region, Region2, Region3, Season,
                              Nano_Silica_Rate, Nano_Carbon_Rate, Nano_Nitrate_Rate, Nano_NP_Rate)

#PCA_df_totals <- data %>% select(Nano_Silica_Rate, Nano_Carbon_Rate, Nano_Nitrate_Rate, Nano_NP_Rate, 
                              #Nitrate, Silicate, SST, MLD, Sigma, PAR, DO, Bottom_depth)

PCA_df_SCC <- data %>% select(Percent_Nano_Silica, Percent_Nano_Carbon, Percent_Nano_Nitrate, Percent_Nano_NP, 
                              Nitrate, Silicate, Temp, MLD, DFS, 
                              Nano_Silica_Rate, Nano_Carbon_Rate, Nano_Nitrate_Rate, Nano_NP_Rate)

PCA_df_SCC2 <- data %>% select(Percent_Nano_Silica, Percent_Nano_Carbon, Percent_Nano_Nitrate, Percent_Nano_NP, 
                              Nitrate, Silicate, SST, DO, PAR, Salinity, POC)


# Create a function to clean data and input for PCA:

PCA_output <- function(df, df2) {
  
  # Omit NaNs:
  PCA_df <- na.omit(df)
  df2 <- na.omit(df2)
  
  # Create the Principal components:
  pca <- prcomp(PCA_df, scale = TRUE, center = TRUE, scores = TRUE)
  
  
  # Create an aoutplot figure to visualize the PCs:
  plot <- autoplot(pca, data = df2, colour = 'Region3', loadings = TRUE, 
         loadings.colour = 'blue',loadings.label = TRUE, 
         loadings.label.size = 3, loadings.label.colour = 'black', 
         frame = T, frame.colour = "Region3", frame.type = 't')
  
  PCA_plot <- plot + 
  #ylim(-0.4, 0.6) +
  xlim(-0.4, 0.8) +
  #guides(fill = FALSE) +
  #scale_color_manual(values = c(8, 9, 3, 6, 5, 13)) +
  theme_pubr()
  
  return(PCA_plot)
  
}


# Small cell contributions PCA:
PCA_output(PCA_df_SCC2, df_SCC)

PCA_output(df_totals, df_totals_cats)

PCA_output(df_sfs, df_sfs_cats)


# Save the figure:
pdf(file = "Rates_PCA.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 6) # The height of the plot in inches

PCA_output(df_totals, df_totals_cats)

dev.off()

```
